Article results

### 2022-07-11 analysis/article-results.Rmd
Crappy attempts with geom_smooth

(Naniar/JointAI summaries/VIM)

@@

knitr::kable(data.frame(mod_coefs))
```

@@

```{r dev-calib}
calplot_MI(
  imps_long = imps_all %>% 
    filter(imps_label == "imps_assess" & dataset == "develop"),
  model = mod_develop_1se, #validation_dev_lambdamin$model_fit,
  impdat_ind = ".imp",
  model_formula = model_formula,
  height_hist = 0.2,
  startpos_hist = -0.25,
  knots = 5,
  n_bins = 100,
  xlim = c(0, 0.85),
  ylim = c(-0.25, 1)
) +
  theme_minimal()
```

@@

# 1 se test
folds <- assign_crossval_folds(y = unique(imps_develop[[".id"]]), n_folds = 10)
cross_val <- glmnet::cv.glmnet(
    x = subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`),
    y = imps_develop$Mc_FailedFemoralApproach,
    wts = imps_develop$wts,
    foldid = folds[match(imps_develop[[".id"]], folds[["y"]]), "fold"],
    family = "binomial",
    alpha = 0
  )
plot(cross_val)

folds <- assign_crossval_folds(y = unique(imps_all[[".id"]]), n_folds = 10)

cross_val <- glmnet::cv.glmnet(
    x = subset(x = stats::model.matrix(model_formula, data = imps_all), select = -`(Intercept)`),
    y = imps_all$Mc_FailedFemoralApproach,
    wts = imps_all$wts,
    foldid = folds[match(imps_all[[".id"]], folds[["y"]]), "fold"],
    family = "binomial",
    alpha = 0
  )
cross_val$lambda
plot(cross_val)

mod_develop_1se <- run_stacked_glmnet(
    x = subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`),
    y = imps_develop$Mc_FailedFemoralApproach,
    wts = imps_develop$wts,
    foldid = folds[match(imps_develop[[".id"]], folds[["y"]]), "fold"],
    lambda_choice = "1se",
    alpha = 0
  )

# Compute linear predictor
lp_valid <- drop(predict(mod_develop_1se, newx = X_valid))
df_calplot_valid <- cbind.data.frame(lp_valid, imps_valid)

@@

calplot_MI(
  imps_long = imps_valid,
  model = mod_develop_1se, #validation_dev_lambdamin$model_fit,
  impdat_ind = ".imp",
  model_formula = model_formula,
  height_hist = 0.2,
  startpos_hist = -0.25,
  knots = 5,
  n_bins = 200,
  xlim = c(0, 0.85),
  ylim = c(-0.25, 1)
) +
  theme_minimal()
```

Try intermediate updating with logreg recalibration

```{r updating}
# 
mod_recal <- split(df_calplot_valid, ~ .imp) |> 
  map(.f = ~ glm(Mc_FailedFemoralApproach ~ lp_valid, family = binomial, data = .x)) |> 
  mice::pool() |> 
  summary()
coefs_mod_recal <- mod_recal$estimate
new_coefs <- drop(coefs_mod_recal[2] * mod_coefs[-1])
new_coefs <- c(mod_coefs[1] + coefs_mod_recal[1], new_coefs) #new_coefs[1] - 0.1059395 #+ coefs_mod_recal[1]
new_lp <- drop((new_coefs) %*% t(cbind(1, X_valid)))

library(Hmisc)
library(rms)

dd <- datadist(imps_valid %>% select(all_of(c(response_var, candidate_predictors)))); 
options(datadist = "dd")

CalibrationCurves::val.prob.ci.2(
  y = (as.numeric(imps_valid$Mc_FailedFemoralApproach) - 1)[1:30000],
  logit = (lp_valid)[1:30000], 
  smooth = FALSE,
  CL.smooth	= FALSE,
  CL.BT	= FALSE,
  logistic.cal = TRUE,
  #g = 10,
  dostats = FALSE
)

df_calplot <- cbind.data.frame("lp_recal" = new_lp, imps_valid)
knots <- 4
#smooth_form <- reformulate(response = response_var, termlabels = paste0("splines::ns(lp, ", knots, ")"))
mod_cal_stacked <- glm(Mc_FailedFemoralApproach ~ ns(lp_recal, 4), data = df_calplot, family = binomial())
stacked_df <- data.frame(
    "predicted" = seq(0.01, 1, by = 0.01),
    "observed" = predict(mod_cal_stacked, 
                         newdata = data.frame("lp_recal" = seq(0.01, 1, by = 0.01)), 
                         type = "response")
)
plot(stacked_df$predicted, stacked_df$observed, xlim = c(0, 1), ylim = c(0, 1))

# Try other

# Try first all in one imp dataset
imp1 <- subset(df_calplot_valid, .imp == 1)

# See https://darrendahly.github.io/post/homr/
m1 <- glm(Mc_FailedFemoralApproach ~ offset(lp_valid), data = imp1, family = binomial())
m2 <- glm(Mc_FailedFemoralApproach ~ lp_valid, data = imp1, family = binomial())
imp1$pred_m1 <- predict(m1, type = "response")
imp1$pred_m2 <- predict(m2, type = "response")

ggplot(data = imp1) +
  geom_smooth(
    aes(x = pred_m1, y = as.numeric(Mc_FailedFemoralApproach) - 1), 
                  color = "red", se = FALSE, method = "loess"
  ) +
  geom_smooth(
    aes(x = plogis(lp_valid), y = as.numeric(Mc_FailedFemoralApproach) - 1), 
                  color = "black", se = FALSE, method = "loess"
  ) + 
  geom_smooth(
    aes(x = pred_m2, y = as.numeric(Mc_FailedFemoralApproach) - 1), 
                  color = "blue", se = FALSE, method = "loess"
  ) + 
  geom_abline(intercept = 0, slope = 1) +
  lims(x = c(0, 1), y = c(0, 1))

# Other tests

## Updating to combined cohort

calplot_MI(
  imps_long = imps_combined,
  model = validation_comb_lambda1se$model_fit,
  impdat_ind = ".imp",
  model_formula = model_formula,
  height_hist = 0.2,
  startpos_hist = -0.25,
  knots = 3,
  n_bins = 200,
  xlim = c(0, 0.55),
  ylim = c(-0.25, 0.8)
) +
  theme_minimal()

# Try the 1se on combined:
imputations <- imps_combined
formula <- model_formula
n_folds = 10
wts <- "wts"
lambda_choice <- "1se"

response_var <- all.vars(stats::update(formula, . ~ 1))
  X_orig <- subset(x = stats::model.matrix(formula, data = imputations), select = -`(Intercept)`)
  y_orig <- imputations[[response_var]]
  wts_orig <- imputations[[wts]]

  folds <- assign_crossval_folds(y = unique(imputations[[".id"]]), n_folds = n_folds)
cross_val <- glmnet::cv.glmnet(
    x = X_orig,
    y = y_orig,
    weights = wts_orig,
    foldid = folds[match(imputations[[".id"]], folds[["y"]]), "fold"],
    family = "binomial", 
    alpha = 0
  )

plot(cross_val)

dat_combined |> 
  select(candidate_predictors, dataset) |>  
    #Mc_FailedFemoralApproach, dataset) |> 
  GGally::ggpairs(
    aes(col = dataset, fill = dataset)
  )

  # First apparent performance - no penalty..
mod <-glmnet::glmnet(
    x = X_orig,
    y = y_orig,
    weights = wts_orig,
    family = "binomial",
    #lambda = 0, # unpenalized
    lambda = cross_val$lambda.min,
    alpha = 0, #ridge
    intercept = TRUE
  )

hist(plogis(predict(mod, newx = X_orig)))
max(plogis(predict(mod, newx = X_orig)))
coef(mod)
coef(
  glm(model_formula, data = imps_combined, family = binomial())
)
calplot_MI(
  imps_long = imps_combined,
  model = mod,
  impdat_ind = ".imp",
  model_formula = model_formula,
  height_hist = 0.2,
  startpos_hist = -0.25,
  knots = 5,
  n_bins = 200,
  xlim = c(0, 1),
  ylim = c(-0.25, 1)
) +
  theme_minimal()

# No penalty
modo <- glm(
  model_formula, family = binomial(), data = imps_combined, weights = imps_combined$wts
)
```

@@

>>>>

<!-- (Naniar/JointAI summaries/VIM) -->

@@

# Print penalied model coefficients
knitr::kable(data.frame(mod_coefs))
```

Bootstrap-based internal validation, optimism-corrected performance:

@@

Calibration plot (apparent)

```{r dev-calib}
calplot_MI(
  imps_long = imps_all %>% 
    filter(imps_label == "imps_assess" & dataset == "develop"),
  model = validation_dev_lambdamin$model_fit,
  impdat_ind = ".imp",
  model_formula = model_formula,
  height_hist = 0.2,
  startpos_hist = -0.25,
  knots = 5,
  n_bins = 100,
  xlim = c(0, 0.85),
  ylim = c(-0.25, 1)
) +
  theme_minimal()
```

@@

# Compute linear predictor
lp_valid <- drop(predict(mod_develop, newx = X_valid))
df_calplot_valid <- cbind.data.frame(lp_valid, imps_valid)

@@

# Overestimation of risks
calplot_MI(
  imps_long = imps_valid,
  model = validation_dev_lambdamin$model_fit,
  impdat_ind = ".imp",
  model_formula = model_formula,
  height_hist = 0.2,
  startpos_hist = -0.25,
  knots = 5,
  n_bins = 200,
  xlim = c(0, 0.85),
  ylim = c(-0.25, 1)
) +
  theme_minimal()
```

## Model update 1: logistic recalibration

```{r log-recal}
# 
imps_valid$lp_valid <- lp_valid
mod_recal <- glm(
  Mc_FailedFemoralApproach ~ lp_valid, 
  family = binomial(), 
  data = imps_valid
)
boots_valid <- group_resample(imps_valid, id_var = ".id", 
                              B = 50, 
    boot_id_name = ".id_boot")

boots_cals <- map_dfr(boots_valid, function(boot_dat) {
  mod_recal_boot <- glm(
  Mc_FailedFemoralApproach ~ lp_valid, 
  family = binomial(), 
  data = boot_dat
)
  calibration_intercept_slope(
    imps_valid$Mc_FailedFemoralApproach, predict(mod_recal_boot, newdata = imps_valid)
  )
})
colMeans(boots_cals)

# Calib plot
library(splines)
ggplot() +
  geom_smooth(
    data = cbind.data.frame(
      "predicted" = predict(mod_recal, type = "response"), #wrrooong
      "outcome" = as.numeric(imps_valid$Mc_FailedFemoralApproach) - 1L,
      "imp" = imps_valid$.imp
    ),
    aes(
      x = predicted,
      y = outcome,
      group = imp
    ), method = "glm",
    formula = y ~ ns(x, 5),
    method.args = list(family = "binomial"),
    se = FALSE,
    col = "gray",
    alpha = 0.7
  ) +
  geom_abline(intercept = 0, slope = 1) +
  lims(x = c(0, 1), y = c(0, 1))
```

## Model update 2: combined cohort re-estimation

### Interaction-by-cohort (unpenalized)

(Ewout reference here, article)

@@

### Internal validation combined cohort

# Try simpler version of cal plot
lp_combined <- predict(
  validation_comb_lambdamin$model_fit,
  newx = subset(
    x = model.matrix(model_formula, data = imps_combined), 
    select = -`(Intercept)`
  )
)

df <- cbind.data.frame(
      "predicted" = drop(lp_combined),
      "outcome" = as.numeric(imps_combined$Mc_FailedFemoralApproach) - 1L,
      "imp" = imps_combined$.imp
    )

binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
}

ggplot(data = df, aes(plogis(predicted), outcome)) +
  binomial_smooth(
    aes(x = predicted),
    formula = y ~ splines::ns(x, 3),
    se = FALSE
  ) #+
  scale_x_continuous(trans = "log", limits = c(0.1, 1))
  geom_smooth(
    ,
    aes(
      x = predicted,
      y = outcome#,
      #group = imp
    ), 
    method = "glm",
    #formula = y ~ splines::ns(x, 3),
    method.args = list(family = "binomial"),
    se = FALSE,
    col = "gray",
    alpha = 0.7
  ) +
  geom_abline(intercept = 0, slope = 1) +
  lims(x = c(0, 1), y = c(0, 1))
